A User-Extensible and Adaptable Parser Architecture

### 2008-08-22 doc/sgai-2008.tex
Corrections and improvements from proofreading.

\newacronym{PDA}{Pushdown Automata}{}
\newacronym{FA}{Finite Automata}{}
\newacronym{CFL}{Context-Free Languages}{}
\newacronym{SMTP}{Simple Mail Transfer Protocol}{}
\newacronym{ATN}{Augmented Transition Networks}{}

@@

The architecture described herein was developed as part of a larger project
to improve anti-spam techniques by analysing the performance of the set of
filters currently in use, optimising the order and membership of the set
based on that analysis, and developing supplemental filters where
deficiencies are identified.  Most anti-spam techniques are
content-based~\cite{a-plan-for-spam, word-stemming, relaxed-online-svms}
which requires accepting the mail in the first place, but rejecting mail
during the delivery attempt is preferable: senders of non-spam mail that is
mistakenly rejected will receive an immediate non-delivery notice; load is
reduced on the accepting mail server (allowing more intensive content-based
techniques to be used); users have less mail to wade through.  Improving
the performance of anti-spam techniques which are applied when mail is
being transferred via \SMTP{}\footnotemark{} is the goal of this project,
by providing a platform for reasoning about anti-spam filters.  The
approach chosen to measure performance is to analyse the log files produced
by the \SMTP{} server in use, Postfix~\cite{postfix}, rather than modifying
it to generate statistics: this approach improves the chances of other
sites testing and utilising the software.  The need arose for a parser
capable of dealing with the great number and variety of log lines produced
by Postfix: the parser must be designed so that adding support for parsing
new inputs is a simple task, as the log lines to be parsed will change over
time.  The variety in log lines occurs for several
reasons:\footnotetext{Simple Mail Transfer Protocol transfers mail across
the Internet from the sender to one or more recipients.  It is a simple,
human readable, plain text protocol, making it quite easy to test and debug
problems with it.  The original protocol definition is
RFC~821~\cite{RFC821}, updated in RFC~2821~\cite{RFC2821}.}

@@

    \item External resources utilised by Postfix (e.g.\ policy
        servers~\cite{policy-servers} or DNSBLs) can also change their
        messages, sometimes without warning.

@@

The solution developed is conceptually simple: provide a few generic
functions (\textit{actions\/}), each capable of dealing with an entire
category of log lines (e.g.\ rejecting a mail), accompanied by a multitude
of precise patterns (\textit{rules\/}), each of which matches all log lines
of a specific type and only that type (e.g.\ rejection by a specific
DNSBL).  It is an accepted standard to separate the parsing procedure from
the declarative grammar it operates with; part of the novelty here is in
the way that the grammar is itself partially procedural (each action is a
separate procedure).  This architecture is ideally suited to parsing inputs
where the input is not fully understood or does not conform to a fixed
grammar: the architecture warns about unparsed inputs and other errors, but
continues on parsing as best it can, allowing the developer of a new parser
to decide which deficiencies are most important and require attention
first, rather than being forced to fix the first error which arises.

@@

For each input the framework tries each rule in turn until it finds one
which matches the input, then invokes the action specified by that rule.

@@

\begin{table}[hbtp]
    \caption{Similarities with ATN}\label{Similarities with ATN}
    % The \smallskip{} below stop the text hitting the lines.
    \begin{tabular}[]{lll}
        \tabletopline{}%
        ATN           & Architecture    & Similarity                                \\
        \tablemiddleline{}%
        Networks      & Framework       & Determines the sequence of transitions or \\
                      &                 & actions which constitutes a valid input.  \\
        Transitions   & Actions         & Saves data and imposes conditions the     \\
                      &                 & input must meet to be considered valid.   \\
        Abbreviations & Rules           & Responsible for classifying input.        \\
        \tablebottomline{}%
    \end{tabular}
\end{table}

@@

\begin{verbatim}
for each line in the input files: 
    for each rule defined by the user: 
        if this rule matches the input line:
            perform the action specified by the rule
            skip the remaining rules
            process the next input line
    warn the user that the input line was not parsed
\end{verbatim}

@@

Each action contains the code required to deal with a particular category
of input, e.g.\ rejections.  The actions are parser-specific: each parser
author will need to write the required actions from scratch, unless
extending an existing parser.  In the Postfix log parser developed for this
project there are \numberOFactions{} actions and \numberOFrules{} rules,
with a very uneven distribution of rules to actions as shown in
\graphref{Distribution of rules per action}.  It is expected that parsers
utilising this architecture will have a high ratio of rules to actions,
hopefully resulting in simpler rules and clearer distinctions between the
inputs parsed by different rules.
\showgraph{build/graph-action-distribution}{Distribution of rules per
action}{Distribution of rules per action} Unsurprisingly the action with
the highest number of associated rules is \texttt{DELIVERY\_REJECTED}, the
action which handles Postfix rejecting a delivery attempt; it is followed
by \texttt{SAVE\_DATA}, the action responsible for handling informative
inputs: these are log lines that supplement the data gathered from other
log lines, but that do not by themselves represent interesting events.  The
third most common action is, perhaps surprisingly, \texttt{UNINTERESTING}:
this action does nothing when executed, allowing uninteresting log lines to
be parsed without causing any effects (it does not imply that the input is
ungrammatical or unparsed).  Generally rules specifying the
\texttt{UNINTERESTING} action are rules to parse log lines which are not
associated with a specific mail, e.g.\ notices about configuration files
changing.  The framework warns about every unparsed input, to alert the
user that they need to alter or extend their ruleset; the Postfix log
parser successfully parses every log line in the \numberOFlogFILESall{} log
files it is currently being tested with.  The remaining actions only have
one or two associated rules: in some cases this is because there will only
ever be one log line variant for that action, e.g.\ all log lines showing
that a remote client has connected are matched by a single rule and handled
by the \texttt{CONNECT} action; in other cases the action is required to
address a deficiency in the log files, or a complication which arises
during parsing.  Taking the \texttt{CONNECT} action as an example, it
creates a new data structure in memory for each client connecting, and
saving the data extracted by the rule; this data will subsequently be
entered into the database.  If a data structure already exists for that
connection it is treated as a symptom of a bug, and the action issues a
warning containing the full contents of the existing data structure, plus
the log line which has just been parsed.

@@

During development of the Postfix log parser it became apparent that in
addition to the obvious variety in log lines, there were a large number of
complications to be overcome (covered in detail in the thesis).  Some were
the result of deficiencies in Postfix's logging (some of which were
improved on or resolved by later versions of Postfix); others were due to
the vagaries of process scheduling, client behaviour, and administrative
actions.  All were successfully accommodated in the Postfix log parser:
adding new actions was enough to overcome several of the complications;
others required modifications to a single existing action to work around
the difficulties; the remainder were resolved by adapting existing actions
to cooperate and exchange extra data, changing their behaviour as
appropriate.

Actions may return a modified input line which will be parsed as if read
from the input stream, allowing for a simplified version of cascaded
parsing~\cite{cascaded-parsing}.  This is a powerful facility allowing a
number of rules and actions to parse a single input line.

@@

Rules categorise inputs, specifying both the regex to match against each
input and the action to invoke when the match is successful.  The Postfix
log parser stores its rules in the same SQL database the results are stored
in, removing any doubt as to which set of rules was used to produce a set
of results; other implementations are free to store the rules in whatever
fashion suits their needs.  The framework requires each rule to have
\texttt{action} and \texttt{regex} attributes; each implementation is free
to add any additional attributes it requires.  The Postfix log parser adds
several attributes to optimise rule ordering (\sectionref{Rule ordering}),
to restrict which log lines each rule can be matched against, and to
describe each rule.  Data is automatically extracted and saved from the log
line based on keywords in the rule's regex, but there is also a facility to
specify extra data to be saved.

@@

The architecture does not try to detect overlapping rules: that
responsibility is left to the author of the rules.  Unintentionally
overlapping rules lead to inconsistent parsing and data extraction because
the first matching rule wins, and the order in which rules are tried
against each input may change between parser invocations.  Overlapping
rules are frequently a requirement, allowing a more specific rule to match
some inputs and a more general rule to match the majority, e.g.\ separating
\SMTP{} delivery to specific sites from \SMTP{} delivery to the rest of the
world.  This can simplify both the general rule and the more specific rule;
additionally rules from different sources can be combined with a minimum of
prior cooperation or modification required.  Overlapping rules should have
a priority attribute to specify their relative ordering; negative
priorities may be useful for catchall rules.

Storing the rules separately to the parser allows external tools to be
written to detect overlapping rules.  Traditional regexes are equivalent in
computational power to \FA{}; it is possible to convert between regexes and
\FA{}, so regex overlap could be detected by finding a non-empty
intersection of two \FA{}\@.  One formula for \FA{} intersection is given
in for example~\cite{intersection-of-NFA-using-Z}: $FA1 \cap{} FA2 =
\overline{(\overline{FA1} \cup{} \overline{FA2})}$.  Perl 5.10 regexes are
more powerful than traditional regexes, e.g.\ it is possible to match
correctly balanced brackets nested to an arbitrary depth:
\newline{}\tab{}\verb!/^[^<>]*(<(?:(?>[^<>]+)|(?1))*>)[^<>]*$/!\newline{}
Perl 5.10 regexes can maintain an arbitrary state stack and are thus
equivalent in computational power to \PDA{}, so overlap detection may
require calculating the intersection of two \PDA{}\@.  The results of
\PDA{} intersection may be intractable in some cases: the intersection of
two \CFL{} (equivalent in computational power to \PDA{}) is not closed,
i.e.\ the resulting language cannot in all cases be parsed by a \CFL{}\@.
A simple example is:
$a^{*}b^{n}c^{n}~\cap~a^{n}b^{n}c^{*}~\rightarrow~a^{n}b^{n}c^{n}$.

@@

It is possible to define pathological regexes, which fall into two main
categories: regexes which match every input line, and regexes which consume
excessive amounts of CPU time during matching.  Defining a regex which
matches every input line is trivial: \verb!/^/! matches the start of every
input line.  In general a regex will not consume much CPU time when the
match is successful; usually CPU time is consumed by a regex with a lot of
alteration and variable quantifiers failing to match.  This topic is beyond
the scope of this paper, see~\cite{mastering-regular-expressions} for
in-depth discussion.

@@

    \item [Partially context-sensitive actions and cascaded parsing:]
        Actions can consult the results (or lack of results) of previous
        actions when executing, providing some context sensitivity.
        Actions can not inspect past or future inputs, however actions
        can return a modified input line that will be parsed as if read
        from the input stream, allowing for a simplified version of
        cascaded parsing~\cite{cascaded-parsing}.

@@

Parsing efficiency is an obvious concern when the Postfix log parser
routinely needs to parse large log files.  The mail server which generated
the log files used in testing the Postfix log parser accepts approximately
10,000 mails for 700 users per day.  Median log file size is 50MB,
containing 285,000 log lines --- large scale mail servers would have much
larger log files.  When generating the timing data used in this section,
\numberOFlogFILES{} log files (totaling 10.08 GB, \numberOFlogLINEShuman{}
log lines) were each parsed 10 times and the parsing times averaged.
Saving results to the database was disabled for the test runs, as the tests
are aimed at measuring the speed of the Postfix log parser rather than the
speed of the database.  The computer used for test runs was a Dell Optiplex
745 described in \tableref{Computer used to generate statistics}; it was
dedicated to the task of gathering statistics from test runs, and was not
used for any other purpose while test runs were ongoing --- any services
not necessary for running the tests were disabled.  Parsing all
\numberOFlogFILES{} log files in one run took
\input{build/include-timing-run-duration}.

@@

An important property of a parser is how parsing time scales relative to
input size: linearly, polynomially, or exponentially?  \Graphref{Parsing
time, file size, and number of lines} shows the parsing time in seconds,
file size in MB, and number of log lines in tens of thousands, for each of
the \numberOFlogFILES{} log files.  All three lines run roughly in
parallel, giving the impression that the algorithm scales linearly with
input size.  This impression is borne out by \graphref{parsing time vs file
size vs number lines factor}: the ratios are tightly banded across the
graph, showing that the algorithm scales linearly.  The ratios increase
(i.e.\ improve) for log files 22 and 62--68 despite their large size; that
unusually large size is due to a mail forwarding loop resulting in a
greatly increased number of mails delivered and log lines generated.

\showgraph{build/graph-input-size-vs-parsing-time}{Parsing time, file size,
and number of lines}{Parsing time, file size, and number of lines}
\showgraph{build/graph-input-size-vs-parsing-time-ratio}{Ratio of number of
log lines and file size to parsing time}{parsing time vs file size vs
number lines factor}

@@

At the time of writing the Postfix log parser has \numberOFrules{}
different rules: the top 10\% match the vast majority of log lines, with
the remaining log lines split across the other 90\% of the rules.  The
distribution resembles a Power Law distribution, as shown in \graphref{rule
hits graph}.  Assuming that the distribution of log lines is reasonably
consistent over time, parser efficiency should benefit from trying more
frequently matching rules before those which match less frequently.  To
test this hypothesis three full test runs were performed with different
rule orderings:

@@

\Graphref{Parsing time relative to shuffled} shows the parsing times of
optimal and reversed orderings as a percentage of the parsing time of
shuffled ordering.  Overall this optimisation provides a modest but
worthwhile reduction in parsing time of approximately 10\%.  Optimal rule
ordering is also beneficial in other ways, as shown in
\sectionref{Architecture Scalability: Number of Rules}.
\showgraph{build/graph-hits}{Hits per rule}{rule hits graph}
\showgraph{build/graph-optimal-and-reverse-vs-shuffle}{Parsing time
relative to shuffled ordering}{Parsing time relative to shuffled}

@@

How any architecture scales as the number of rules increases is important,
but it is particularly important in this architecture because it is
expected that typical parsers will have a large number of rules.  There are
\numberOFrules{} rules in the full Postfix log parser ruleset, whereas the
minimum number of rules required to parse the \numberOFlogFILES{} log files
is \numberOFrulesMINIMUM{}, \numberOFrulesMINIMUMpercentage{} of the full
ruleset.  A second set of statistics was generated using the minimum
ruleset, and the parsing times compared to those generated using the full
ruleset: the percentage parsing time increase when using the full ruleset
instead of the minimal ruleset for optimal, shuffled and reversed orderings
is shown in \graphref{Percentage parsing time increase of maximum ruleset
over minimum ruleset}.  It is clear that the increased number of rules has
a noticeable performance impact with reverse ordering, and a lesser impact
with shuffled ordering.  The optimal ordering shows a mean increase of
\input{build/include-full-ruleset-vs-minimum-ruleset} in parsing time for a
\numberOFrulesMAXIMUMpercentage{} increase in the number of rules.  These
results show that the architecture scales very well as the number of rules
increases, and that optimally ordering the rules is an important
optimisation enabling this.
\showgraph{build/graph-full-ruleset-vs-minimum-ruleset}{Percentage parsing
time increase of maximum ruleset over minimum ruleset}{Percentage parsing
time increase of maximum ruleset over minimum ruleset}

@@

Correct parsing of log lines must be measured first.  Warnings are issued
for any log lines which are not parsed; no such warnings are issued while
parsing the \numberOFlogFILES{} log files, so it can be safely concluded
that there are zero false negatives.  False positives are harder to
quantify: manually verifying that the correct rule parsed each of the
\numberOFlogLINES{} log lines is infeasible, making it impossible to
quantify the false positive rate.  A random sample of 6039 log lines
(0.00994\% of the total) was parsed and the results manually verified by
inspection to ensure that the correct rule parsed each log line.  The
sample results contained zero false positives, and this check has been
automated to ensure continued accuracy.

The proportion of mail delivery attempts correctly understood is much more
difficult to determine accurately than log lines correctly parsed.  The
implementation can dump its state tables in a human readable form;
examining these tables with reference to the log files and database is the
best way to detect mails which were not handled properly.  The
implementation issues warnings when it detects any errors, which should
alert the user to a problem.  There should be few or no warnings during
parsing, and when parsing is finished the state table should only contain
entries for mails starting before or ending after the log file.  A second
sample of 6000 log lines was parsed with all debugging options enabled,
resulting in 167,448 lines of output.  All 167,448 lines were examined in
conjunction with the input log file and a dump of the resulting database,
verifying that for each of the log lines the implementation performed
correctly.  The implementation produced 4 warnings about deficiencies in
the log segment, 10 mails remaining in the state tables, and 1625 correct
entries in the database.  No error messages were produced, therefore there
were no false negatives.  Given the evidence detailed above, the authors
are confident that the false positive rate when understanding a mail
delivery attempt is zero.

@@

This architecture's greatest strength is the ease with which it can be
adapted to deal with new requirements and inputs.  Parsing a variation of
an existing input is a trivial task: simply modify an existing rule or add
a new rule with an appropriate regex and the task is complete; knowledge of
the framework is not required.  Parsing a new category of input is achieved
by writing a new action and appropriate rules; quite often the action will
not need to interact with existing actions, but when interaction is
required the framework provides the necessary facilities.  The decoupling
of rules from actions allows different sets of rules to be used with the
same actions, e.g.\ a parser might have actions to process versions one and
two of a file format; the parser will then parse version one, version two,
or both versions, depending on the ruleset selected.  The architecture
makes it possible to apply commonly used programming techniques such as
object orientation, inheritance, composition, delegation, and roles when
designing and implementing a parser, simplifying the process of working
within a team with shared responsibility for a parser, or when developing
and testing additional functionality.  This architecture is ideally suited
to parsing inputs where the input is not fully understood or does not
follow a fixed grammar: the architecture warns about unparsed inputs and
other errors, but continues on parsing as best it can, allowing the
developer of a new parser to decide which deficiencies are most important
and require attention first, rather than being forced to fix the first
error which arises.

>>>>

\newacronym{PDA}{Pushdown Automata}{}
\newacronym{FA}{Finite Automata}{}
\newacronym{CFL}{Context-Free Languages}{}
\newacronym{SMTP}{Simple Mail Transfer Protocol}{}
\newacronym{ATN}{Augmented Transition Networks}{}
\newacronym{DNSBL}{DNS Black List}{}

@@

The architecture described herein was developed as part of a larger project
to improve anti-spam techniques by analysing the performance of the set of
filters currently in use, optimising the order and membership of the set
based on that analysis, and developing supplemental filters where
deficiencies are identified.  Most anti-spam techniques are
content-based~\cite{a-plan-for-spam, word-stemming, relaxed-online-svms}
which requires accepting the mail in the first place to determine if it is
spam, but rejecting mail during the delivery attempt is preferable: senders
of non-spam mail that is mistakenly rejected will receive an immediate
non-delivery notice; load is reduced on the accepting mail server (allowing
more intensive content-based techniques to be used); users have less mail
to wade through.  Improving the performance of anti-spam techniques which
are applied when mail is being transferred via \SMTP{}\footnotemark{} is
the goal of this project, by providing a platform for reasoning about
anti-spam filters.  The approach chosen to measure performance is to
analyse the log files produced by the \SMTP{} server in use,
Postfix~\cite{postfix}, rather than modifying it to generate statistics:
this approach improves the chances of other sites testing and utilising the
software.  The need arose for a parser capable of dealing with the great
number and variety of log lines produced by Postfix: the parser must be
designed so that adding support for parsing new inputs is a simple task, as
the log lines to be parsed will change over time.  The variety in log lines
occurs for several reasons:\footnotetext{Simple Mail Transfer Protocol
transfers mail across the Internet from the sender to one or more
recipients.  It is a simple, human readable, plain text protocol, making it
quite easy to test and debug problems with it.  The original protocol
definition is RFC~821~\cite{RFC821}, updated in RFC~2821~\cite{RFC2821}.}

@@

    \item External resources utilised by Postfix (e.g.\ \DNSBL{} or policy
        servers~\cite{policy-servers}) can also change their messages,
        sometimes without warning.

@@

The solution developed is conceptually simple: provide a few generic
functions (\textit{actions\/}), each capable of dealing with an entire
category of log lines (e.g.\ rejecting a mail), accompanied by a multitude
of precise patterns (\textit{rules\/}), each of which matches all log lines
of a specific type and only that type (e.g.\ rejection by a specific
\DNSBL{}).  It is an accepted standard to separate the parsing procedure
from the declarative grammar it operates with; part of the novelty here is
in the way that the grammar is itself partially procedural (each action is
a separate procedure).  This architecture is ideally suited to parsing
inputs where the input is not fully understood or does not conform to a
fixed grammar: the architecture warns about unparsed inputs and other
errors, but continues on parsing as best it can, allowing the developer of
a new parser to decide which deficiencies are most important and require
attention first, rather than being forced to fix the first error which
arises.

@@

For each input the framework tries each rule in turn until it finds a rule
which matches the input, then invokes the action specified by that rule.

@@

\begin{table}[hbtp]
    \caption{Similarities with ATN}\label{Similarities with ATN}
    % The \smallskip{} below stop the text hitting the lines.
    \begin{tabular}[]{lll}
        \tabletopline{}%
        ATN           & Parser Architecture & Similarity                                \\
        \tablemiddleline{}%
        Networks      & Framework           & Determines the sequence of transitions or \\
                      &                     & actions which constitutes a valid input.  \\
        Transitions   & Actions             & Saves data and imposes conditions the     \\
                      &                     & input must meet to be considered valid.   \\
        Abbreviations & Rules               & Responsible for classifying input.        \\
        \tablebottomline{}%
    \end{tabular}
\end{table}

@@

\begin{verbatim}
for each input:
    for each rule defined by the user: 
        if this rule matches the input:
            perform the action specified by the rule
            skip the remaining rules
            process the next input
    warn the user that the input was not parsed
\end{verbatim}

@@

Each action contains the code required to deal with a particular category
of input, e.g.\ rejections.  The actions are parser-specific: each parser
author will need to write the required actions from scratch, unless
extending an existing parser.  It is expected that parsers utilising this
architecture will have a high ratio of rules to actions, hopefully
resulting in simpler rules and clearer distinctions between the inputs
parsed by different rules.  In the Postfix log parser developed for this
project there are \numberOFactions{} actions and \numberOFrules{} rules,
with a very uneven distribution of rules to actions as shown in
\graphref{Distribution of rules per action}.
\showgraph{build/graph-action-distribution}{Distribution of rules per
action}{Distribution of rules per action} Unsurprisingly the action with
the highest number of associated rules is \texttt{DELIVERY\_REJECTED}, the
action which handles Postfix rejecting a mail delivery attempt; it is
followed by \texttt{SAVE\_DATA}, the action responsible for handling
informative log lines, which supplement the data gathered from other log
lines but do not by themselves represent interesting events.  The third
most common action is, perhaps surprisingly, \texttt{UNINTERESTING}: this
action does nothing when executed, allowing uninteresting log lines to be
parsed without causing any effects (it does not imply that the input is
ungrammatical or unparsed).  Generally rules specifying the
\texttt{UNINTERESTING} action are rules to parse log lines which are not
associated with a specific mail, e.g.\ notices about configuration files
changing.  The remaining actions only have one or two associated rules: in
some cases this is because there will only ever be one log line variant for
that action, e.g.\ all log lines showing that a remote client has connected
are matched by a single rule and handled by the \texttt{CONNECT} action; in
other cases the action is required to address a deficiency in the log
files, or a complication which arises during parsing.  Taking the
\texttt{CONNECT} action as an example, it creates a new data structure in
memory for each client connecting, and saves the data extracted by the
rule; this data will subsequently be entered into the database.  If a data
structure already exists for that connection it is treated as a symptom of
a bug, and the action issues a warning containing the full contents of the
existing data structure, plus the log line which has just been parsed.

@@

During development of the Postfix log parser it became apparent that in
addition to the obvious variety in log lines, there were a large number of
complications to be overcome.  Some were the result of deficiencies in
Postfix's logging (some of which were improved on or resolved by later
versions of Postfix); others were due to the vagaries of process
scheduling, client behaviour, and administrative actions.  All were
successfully accommodated in the Postfix log parser: adding new actions was
enough to overcome several of the complications; others required
modifications to a single existing action to work around the difficulties;
the remainder were resolved by adapting existing actions to cooperate and
exchange extra data, changing their behaviour as appropriate.

Actions may return a modified input line which will be parsed as if read
from the input stream, allowing for a simplified version of cascaded
parsing~\cite{cascaded-parsing}.  This is a powerful facility allowing a
number of rules and actions to parse a single input.

@@

Rules categorise inputs, specifying both the regex to match against each
input and the action to invoke when the match is successful.  The Postfix
log parser stores its rules in the same SQL database the results are stored
in, removing any doubt as to which set of rules was used to produce a set
of results; other implementations are free to store the rules in whatever
fashion suits their needs.  The framework warns about every unparsed input,
to alert the user that they need to alter or extend their ruleset; the
Postfix log parser successfully parses every log line in the
\numberOFlogFILESall{} log files it is currently being tested with.  The
framework requires each rule to have \texttt{action} and \texttt{regex}
attributes; each implementation is free to add any additional attributes it
requires.  The Postfix log parser adds several attributes to optimise rule
ordering (\sectionref{Rule ordering}), to restrict which log lines each
rule can be matched against, and to describe each rule.  Data is
automatically extracted and saved from the log line based on keywords in
the rule's regex, but there is also a facility to specify extra data to be
saved.

@@

The architecture does not try to detect overlapping rules: that
responsibility is left to the author of the rules.  Unintentionally
overlapping rules lead to inconsistent parsing and data extraction because
the first matching rule wins, and the order in which rules are tried
against each input may change between parser invocations.  Overlapping
rules are frequently a requirement, allowing a more specific rule to match
some inputs and a more general rule to match the remainder, e.g.\
separating \SMTP{} delivery to specific sites from \SMTP{} delivery to the
rest of the world.  This can simplify both the general rule and the more
specific rule; additionally rules from different sources can be combined
with a minimum of prior cooperation or modification required.  Overlapping
rules should have a priority attribute to specify their relative ordering;
negative priorities may be useful for catchall rules.

Storing the rules separately to the parser allows external tools to be
written to detect overlapping rules.  Traditional regexes are equivalent in
computational power to \FA{}, and it is possible to convert between the
two, so regex overlap could be detected by finding a non-empty intersection
of two \FA{}\@.  The standard equation for \FA{} intersection (given for
example in~\cite{intersection-of-NFA-using-Z}) is: $FA1 \cap{} FA2 =
\overline{(\overline{FA1} \cup{} \overline{FA2})}$.  Perl 5.10 regexes are
more powerful than traditional regexes, e.g.\ it is possible to match
correctly balanced brackets nested to an arbitrary depth:
\newline{}\tab{}\verb!/^[^<>]*(<(?:(?>[^<>]+)|(?1))*>)[^<>]*$/!\newline{}
Perl 5.10 regexes can maintain an arbitrary state stack and are thus
equivalent in computational power to \PDA{} or \CFL{}, so detecting overlap
may require calculating the intersection of two \PDA{} or \CFL{}\@.  The
intersection of two \CFL{} is not closed, i.e.\ the resulting language
cannot in all cases be parsed by a \CFL{}, so intersection may be
intractable in some cases e.g.:
$a^{*}b^{n}c^{n}~\cap~a^{n}b^{n}c^{*}~\rightarrow~a^{n}b^{n}c^{n}$.

@@

It is possible to define pathological regexes, which fall into two main
categories: regexes which match every input, and regexes which consume
excessive amounts of CPU time during matching.  Defining a regex which
matches every input is trivial: \verb!/^/! matches the start of every
input.  In general a regex will not consume much CPU time when the match is
successful; usually CPU time is consumed when a regex with a lot of
alteration and variable quantifiers fails to match.  This topic is beyond
the scope of this paper, see~\cite{mastering-regular-expressions} for
in-depth discussion.

@@

    \item [Partially context-sensitive actions and cascaded parsing:]
        Actions can consult the results (or lack of results) of previous
        actions during execution, providing some context sensitivity.
        Actions can not inspect past or future inputs, however actions can
        return a modified input that will be parsed as if read from the
        input stream, allowing for a simplified version of cascaded
        parsing~\cite{cascaded-parsing}.

@@

Parsing efficiency is an obvious concern when the Postfix log parser
routinely needs to parse large log files.  The mail server which generated
the log files used in testing the Postfix log parser accepts approximately
10,000 mails for 700 users per day.  Median log file size is 50MB,
containing 285,000 log lines --- large scale mail servers would have much
larger log files.  When generating the timing data used in this section,
\numberOFlogFILES{} log files (totaling 10.08 GB, \numberOFlogLINEShuman{}
log lines) were each parsed 10 times and the parsing times averaged.
Saving results to the database was disabled for the test runs, as the tests
are aimed at measuring the speed of the Postfix log parser rather than the
speed of the database.  The computer used for test runs is a Dell Optiplex
745 described in \tableref{Computer used to generate statistics}; it was
dedicated to the task of gathering statistics from test runs, and was not
used for any other purpose while test runs were ongoing --- any services
not necessary for running the tests were disabled.  Parsing all
\numberOFlogFILES{} log files in one run took
\input{build/include-timing-run-duration}.

@@

An important property of a parser is how parsing time scales relative to
input size: linearly, polynomially, or exponentially?  \Graphref{Parsing
time, file size, and number of log lines} shows the parsing time in
seconds, file size in MB, and number of log lines in tens of thousands, for
each of the \numberOFlogFILES{} log files.  All three lines on the graph
run roughly in parallel, giving the impression that the algorithm scales
linearly with input size.  This impression is borne out by
\graphref{parsing time vs file size vs number lines factor}: the ratios are
tightly banded across the graph, showing that the algorithm scales
linearly.  The ratios increase (i.e.\ improve) for log files 22 and 62--68
despite their large size; that unusually large size is due to a mail
forwarding loop resulting in a greatly increased number of mails delivered
and log lines generated.

\showgraph{build/graph-input-size-vs-parsing-time}{Parsing time, file size,
and number of log lines}{Parsing time, file size, and number of log lines}
\showgraph{build/graph-input-size-vs-parsing-time-ratio}{Ratio of number of
log lines and file size to parsing time}{parsing time vs file size vs
number lines factor}

@@

At the time of writing the Postfix log parser has \numberOFrules{}
different rules: as shown in \graphref{rule hits graph} 10\% of the rules
match the vast majority of log lines, with the remaining log lines split
across the other 90\% of the rules, similar to a Power Law distribution.
Assuming that the distribution of log lines is reasonably consistent over
time, parser efficiency should benefit from trying more frequently matching
rules before those which match less frequently.  To test this hypothesis
three full test runs were performed with different rule orderings:

\begin{description}[optimal]

@@

\Graphref{Parsing time relative to shuffled} shows the parsing times of
optimal and reversed orderings as a percentage of the parsing time of
shuffled ordering.  Overall this optimisation provides a modest but
worthwhile reduction in parsing time of approximately 10\%.  Optimal rule
ordering is also beneficial in other ways, as described in
\sectionref{Architecture Scalability: Number of Rules}.
\showgraph{build/graph-hits}{Number of log lines matched by each rule}{rule
hits graph} \showgraph{build/graph-optimal-and-reverse-vs-shuffle}{Parsing
time relative to shuffled ordering}{Parsing time relative to shuffled}

@@

How any architecture scales as the number of rules increases is important,
but it is particularly important for this architecture because it is
expected that typical parsers will have a large number of rules.  There are
\numberOFrules{} rules in the full Postfix log parser ruleset, whereas the
minimum number of rules required to parse the \numberOFlogFILES{} log files
is \numberOFrulesMINIMUM{}, \numberOFrulesMINIMUMpercentage{} of the full
ruleset.  A second set of statistics was generated using the minimum
ruleset, and the parsing times compared to those generated using the full
ruleset: the percentage parsing time increase when using the full ruleset
instead of the minimal ruleset for optimal, shuffled and reversed orderings
is shown in \graphref{Percentage parsing time increase of maximum ruleset
over minimum ruleset}.  It is clear that the increased number of rules has
a noticeable performance impact with reverse ordering, and a lesser impact
with shuffled ordering.  The optimal ordering shows a mean increase of
\input{build/include-full-ruleset-vs-minimum-ruleset} in parsing time for a
\numberOFrulesMAXIMUMpercentage{} increase in the number of rules.  These
results show that the architecture scales very well as the number of rules
increases, and that optimally ordering the rules is an important
optimisation enabling this.
\showgraph{build/graph-full-ruleset-vs-minimum-ruleset}{Percentage parsing
time increase of maximum ruleset over minimum ruleset}{Percentage parsing
time increase of maximum ruleset over minimum ruleset}

@@

Correct parsing of log lines must be measured first.  Warnings are issued
for any log lines which are not parsed; no such warnings are issued while
parsing the \numberOFlogFILES{} log files, so it can be safely concluded
that there are zero false negatives.  False positives are harder to
quantify: manually verifying that the correct rule parsed each of the
\numberOFlogLINES{} log lines is infeasible.  A random sample of 6,039 log
lines (0.00994\% of \numberOFlogLINES{}) was parsed and the results
manually verified by inspection to ensure that the correct rule parsed each
log line.  The sample results contained zero false positives, and this
check has been automated to ensure continued accuracy.  The authors are
confident that zero false positives occur when parsing the
\numberOFlogFILES{} log files.

The proportion of mail delivery attempts correctly understood is much more
difficult to determine accurately than the proportion of log lines
correctly parsed.  The implementation can dump its state tables in a human
readable form; examining these tables with reference to the log files and
database is the best way to detect mails which were not handled properly.
The implementation issues warnings when it detects any errors or
discrepancies, which should alert the user to a problem.  There should be
few or no warnings during parsing, and when parsing is finished the state
table should only contain entries for mails starting before or ending after
the log file.  A second sample of 6000 log lines was parsed with all
debugging options enabled, resulting in 167,448 lines of output.  All
167,448 lines were examined in conjunction with the log segment and a dump
of the resulting database, verifying that for each of the log lines the
implementation performed correctly.  The implementation produced 4 warnings
about deficiencies in the log segment, 10 mails correctly remaining in the
state tables, and 1625 correct entries in the database.  No error messages
were produced, therefore there were no false negatives.  Given the evidence
detailed above, the authors are confident that the false positive rate when
understanding a mail delivery attempt is zero.

@@

This architecture's greatest strength is the ease with which it can be
adapted to deal with new requirements and inputs.  Parsing a variation of
an existing input is a trivial task: simply modify an existing rule or add
a new rule with an appropriate regex and the task is complete.  Parsing a
new category of input is achieved by writing a new action and appropriate
rules; quite often the new action will not need to interact with existing
actions, but when interaction is required the framework provides the
necessary facilities.  The decoupling of rules from actions allows
different sets of rules to be used with the same actions, e.g.\ a parser
might have actions to process versions one and two of a file format; the
parser will then parse version one, or version two, or both versions,
depending on the ruleset selected.  The architecture makes it possible to
apply commonly used programming techniques such as object orientation,
inheritance, composition, delegation, and roles when designing and
implementing a parser, simplifying the process of working within a team
with shared responsibility for a parser, or when developing and testing
additional functionality.  This architecture is ideally suited to parsing
inputs where the input is not fully understood or does not follow a fixed
grammar: the architecture warns about unparsed inputs and other errors, but
continues on parsing as best it can, allowing the developer of a new parser
to decide which deficiencies are most important and require attention
first, rather than being forced to fix the first error which arises.
